---
title: "R Notebook"
output: html_notebook
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tinytex)
library(ggplot2)
library(dplyr)
library(tidyverse)
```

# Q1 Suppose that we have age data including the following numbers in sorted order.
# age: 13, 15, 16, 16, 19, 20, 20, 21, 22, 22, 25, 25, 25, 25, 30, 33, 33, 35, 35, # 35, 35, 36, 40, 45, 46, 52, 70

# (a)
# Use smoothing by bin means to smooth the above data, using a bin depth of 3

```{r}

age_data_df = c(13, 15, 16, 16, 19, 20, 20, 21, 22, 22, 25, 25, 25, 25, 30, 33, 33, 35, 35, 35, 35, 36, 40, 45, 46, 52, 70)

bin_depth = 3

age_data_bin = split(age_data_df,ceiling(seq_along(age_data_df) / bin_depth))



mean_age_data_bin1 = sapply(age_data_bin[1], mean)
mean_age_data_bin2 = sapply(age_data_bin[2], mean)
mean_age_data_bin3 = sapply(age_data_bin[3], mean)
mean_age_data_bin4 = sapply(age_data_bin[4], mean)
mean_age_data_bin5 = sapply(age_data_bin[5], mean)
mean_age_data_bin6 = sapply(age_data_bin[6], mean)
mean_age_data_bin7 = sapply(age_data_bin[7], mean)
mean_age_data_bin8 = sapply(age_data_bin[8], mean)
mean_age_data_bin9 = sapply(age_data_bin[9], mean)
```

# (b)
# Use IQR measure to determine if there are any outliers in this data

```{r}
age_data_median = median(age_data_df)
age_data_median_Q1 = median(age_data_df[1:14])
age_data_median_Q3 = median(age_data_df[15:27])
age_data_IRQ = age_data_median_Q3 - age_data_median_Q1

age_data_upper = age_data_median_Q3 + (1.5 * age_data_IRQ)
age_data_lower = age_data_median_Q1 - (1.5 * age_data_IRQ)

age_data_outliers = age_data_df[age_data_df > 56.5]
age_data_outliers


```

# (c)
# Use min-max normalization to transform the value 35 for age onto the range [0.0, 1.0].

# v2 = ((v - minA) / (maxA - minA)) (new_maxA - new_minA) + new_minA

```{r}

v = 35
minA = 13
maxA = 70
new_minA = 0.0
new_maxA = 1.0

v2 = ((v - minA) / (maxA - minA)) * (new_maxA - new_minA) + new_minA
v2



```

# (d)
# Use z-score normalization to transform the value 35 for age? (you need to compute mean and standard deviation first)

# v_z_score = (v - mean) / standard deviation


```{r}

mean_age_data = mean(age_data_df)
sd_age_data = sd(age_data_df)

v_z_score = (v - mean_age_data) / sd_age_data
v_z_score

```
# (e)
# Use normalization by decimal scaling to transform the value 35 for age

# v' = v / 10j where Where j is the smallest integer such that Max(|ν’|) < 1


```{r}
age_data_df_one_tenth = age_data_df / 10
age_data_df_one_hundreth = age_data_df / 100

v_dec_scale = v / (10^2)
v_dec_scale



```



# Q2. Write a function in your preferred language which can take a data vector and do min-max normalization by transforming
# data onto a desired range. For example, it should be able get the age data above and map it between any two numbers.
# (foo (a, min_new, max_new) where a is an one-dimensional array)


```{r}
min_age = min(age_data_df)
max_age = max(age_data_df)


min_max_norm <- function(a, min_new, max_new)
{

  ((a - min_age) / (max_age - min_age)) * (max_new - min_new) + min_new

}

min_max_norm(35, 0.0, 1.0)
min_max_norm(age_data_df[9], 0.0, 1.0)
min_max_norm(age_data_df[9], 0.3, 2.0)
min_max_norm(age_data_df[15], 0.0, 1.0)
```

# Using information gain on the data in Table 1, do calculations for two levels of a decision tree which decides whether
# a person is senior or junior. 

```{r}




```


